Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-07-22 manuscript_results.Rmd
updated code

```{r}
plot_forecast <- function(i, ii,
                          list_of_states,
                          AUTO_AVERAGE_WEEK1_list,
                          AUTO_AVERAGE_WEEK2_list,
                          AUTO_AVERAGE_WEEK3_list,
                          AUTO_AVERAGE_WEEK4_list) {

  # Define date range for 2-year history
  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
  start_date <- end_date - (7 * 103)

  # Subset observed data
  subset_data <- subset(list_of_states[[ii]],
                        target_end_date >= start_date & target_end_date <= end_date)

  # Build forecast dataframe
  forecast_df <- data.frame(
    Dates = c(
      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
    ),
    Hosp = c(
      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
    ),
    Value = c(
      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
    ),
    Lower = c(
      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
    ),
    Upper = c(
      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
    )
  )

  # Build the plot
  ggplot() +
    geom_point(data = subset_data,
               aes(x = target_end_date, y = cases, color = "Fitting data"),
               size = 2.5) +
    geom_ribbon(data = forecast_df,
                aes(x = Dates, ymin = Lower, ymax = Upper),
                fill = "gray", alpha = 0.4) +
    geom_point(data = forecast_df,
               aes(x = Dates, y = Value, color = "Mean forecast"),
               size = 3) +
    geom_point(data = forecast_df,
               aes(x = Dates, y = Value, color = "Mean forecast"),
               shape = 21, fill = "white", size = 2) +
    geom_point(data = forecast_df,
               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
               shape = 16, size = 2.5) +
    scale_color_manual(name = "Legend",
                       values = c("Fitting data" = "red",
                                  "Mean forecast" = "cyan3",
                                  "Hospitalizations" = "black")) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
         subtitle = "Gray ribbon represents 95% confidence interval",
         x = "", y = "Hospitalizations") +
    theme_minimal() +
    theme(plot.title = element_text(size = 18, face = "bold"),
          plot.subtitle = element_text(size = 14),
          axis.text.x = element_text(angle = 45, hjust = 1))
}

for (ii in 1:48) {
  for (i in 60:70) {
    # Safely check if index exists
    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
      p <- plot_forecast(i, ii,
                         list_of_states,
                         AUTO_AVERAGE_WEEK1_list,
                         AUTO_AVERAGE_WEEK2_list,
                         AUTO_AVERAGE_WEEK3_list,
                         AUTO_AVERAGE_WEEK4_list)
      # Show the plot
      print(p)
    }
  }
}

@@

# Combine your main plot with the custom legend
figure4 <- combined_plot_clean + 
  patchwork::plot_spacer() + 
  patchwork::wrap_elements(legend_only) + 
  plot_layout(widths = c(4, 0.05, 0.4))  # Adjust widths as needed

@@

ERASE

```{r}
# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[2], "ES64_AR_LB" = colors_ar[3],
  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[2], "ES64_EPI_LB" = colors_epi[3],
  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[2], "ES64_TMP_LB" = colors_tmp[3],
  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[2], "ES64_ADJ_LB" = colors_adj[3],
  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[2], "ES64_AVG_LB" = colors_avg[3],

  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
  "AUTO_AVG" = colors_avg[1], "ES27_AVG" = colors_avg[2], "ES64_AVG" = colors_avg[3]
)

# --------- WEEK1 MODELS ------------- #
plot1<- ggplot(W1, aes(x = Best_Result, fill = Best_Result)) +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W1)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")

#ggsave("week1_best_models.jpg", plot1)

# --------- WEEK2 MODELS ------------- #
plot2<-ggplot(W2, aes(x = Best_Result, fill = Best_Result))  +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W2)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")

#ggsave("week2_best_models.jpg", plot2)

# --------- WEEK3 MODELS ------------- #
plot3<-ggplot(W3, aes(x = Best_Result, fill = Best_Result))  +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W3)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")
plot3

#ggsave("week3_best_models.jpg", plot3)

# --------- WEEK4 MODELS ------------- #
plot4<-ggplot(W4, aes(x = Best_Result, fill = Best_Result)) +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W4)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")
plot4

#ggsave("week4_best_models.jpg", plot4)

ERASE

```{r}
# Add a column to indicate the week ahead
W1$Week_Ahead <- "W1"
W2$Week_Ahead <- "W2"
W3$Week_Ahead <- "W3"
W4$Week_Ahead <- "W4"

# Combine all weeks into one dataframe
all_weeks <- bind_rows(W1, W2, W3, W4)

# Create the plot
combined_plot <- ggplot(all_weeks, aes(x = Best_Result, fill = Best_Result)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Best model based on the lowest mean WIS",
       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
       x = "", y = "Number of states",
       fill = "Best model" ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold")) +
  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x")

# Save and display the plot
#ggsave("Fig4.jpg", combined_plot, width =8, height =5)
print(combined_plot)

Now we use a Wilcox test with Holm p-value adjustment for repeated comparisons to evaluate the differences in model performances.

@@

# Create a new column "Model_Type" based on model with and withouth log-back transformation
p_values_and_impr$Model_Type <- ifelse(grepl("_LB$", p_values_and_impr$Model), "LB", "no_LB")

# View the updated dataframe
head(p_values_and_impr)
p_values_and_impr
```
ERASE

#new_long_data <- summary_impr %>%
#  mutate(
#    LogBack = grepl("_LB$", Model),
#    Family = sub("_.*", "", Model),  # take everything before the first underscore
#    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
#  )

#new_long_data <- new_long_data %>%
#  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))

#model <- lm(m ~ Model, data = new_long_data)
#summary(model)

#model
#new_long_data$Model
# Get tidy summary of coefficients
#tidy_model <- broom::tidy(model)

# Filter out intercept for clarity if you want
#tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")

# Plot coefficients with 95% confidence intervals
#tidy_model_no_intercept <- tidy_model_no_intercept %>%
#  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))

# Plot
#ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
#  geom_point() +
#  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
#                    ymax = estimate + std.error * 1.96),
#                width = 0.2) +
#  coord_flip() +  # flip for better readability
#  labs(
#    title = "Effect Size of Model Types on outcome",
#    subtitle =  "Estimates relative to baseline (intercept)",
#    y = "Estimate",
#    x = ""
#  ) +
#  theme_minimal()

@@

ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
```

@@

ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

#ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
```

@@

#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
```

@@

#ggsave(ES_4WEEK, height = 4.5, width = 6.5,  filename="Fig5D.jpg")  
```

@@

#ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
```
Total hospitalizations

>>>>

```{r}
#plot_forecast <- function(i, ii,
#                          list_of_states,
#                          AUTO_AVERAGE_WEEK1_list,
#                          AUTO_AVERAGE_WEEK2_list,
#                          AUTO_AVERAGE_WEEK3_list,
#                          AUTO_AVERAGE_WEEK4_list) {

  # Define date range for 2-year history
#  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
#  start_date <- end_date - (7 * 103)

  # Subset observed data
#  subset_data <- subset(list_of_states[[ii]],
#                        target_end_date >= start_date & target_end_date <= end_date)

  # Build forecast dataframe
#  forecast_df <- data.frame(
#    Dates = c(
#      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
#      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
#      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
#      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
#    ),
#    Hosp = c(
#      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
#      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
#      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
#      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
#    ),
#    Value = c(
#      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
#      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
#      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
#      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
#    ),
#    Lower = c(
#      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
#      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
#      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
#      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
#    ),
#    Upper = c(
#      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
#      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
#      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
#      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
#    )
#  )

  # Build the plot
#  ggplot() +
#    geom_point(data = subset_data,
#               aes(x = target_end_date, y = cases, color = "Fitting data"),
#               size = 2.5) +
#    geom_ribbon(data = forecast_df,
#                aes(x = Dates, ymin = Lower, ymax = Upper),
#                fill = "gray", alpha = 0.4) +
#    geom_point(data = forecast_df,
#               aes(x = Dates, y = Value, color = "Mean forecast"),
#               size = 3) +
#    geom_point(data = forecast_df,
#               aes(x = Dates, y = Value, color = "Mean forecast"),
#               shape = 21, fill = "white", size = 2) +
#    geom_point(data = forecast_df,
#               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
#               shape = 16, size = 2.5) +
#    scale_color_manual(name = "Legend",
#                       values = c("Fitting data" = "red",
#                                  "Mean forecast" = "cyan3",
#                                  "Hospitalizations" = "black")) +
#    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
#    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
#         subtitle = "Gray ribbon represents 95% confidence interval",
#         x = "", y = "Hospitalizations") +
#    theme_minimal() +
#    theme(plot.title = element_text(size = 18, face = "bold"),
#          plot.subtitle = element_text(size = 14),
#          axis.text.x = element_text(angle = 45, hjust = 1))
#}

#for (ii in 1:48) {
#  for (i in 60:70) {
    # Safely check if index exists
#    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
#      p <- plot_forecast(i, ii,
#                         list_of_states,
#                         AUTO_AVERAGE_WEEK1_list,
#                         AUTO_AVERAGE_WEEK2_list,
#                         AUTO_AVERAGE_WEEK3_list,
#                         AUTO_AVERAGE_WEEK4_list)
      # Show the plot
#      print(p)
#    }
#  }
#}

@@

# Combine main plot with the custom legend
figure4 <- combined_plot_clean + 
  patchwork::plot_spacer() + 
  patchwork::wrap_elements(legend_only) + 
  plot_layout(widths = c(4, 0.05, 0.4))

@@

Now we use a Wilcoxon test with Holm p-value adjustment for repeated comparisons to evaluate the differences in model performances.

@@

# Create a new column "Model_Type" based on model with and without log-back transformation
p_values_and_impr$Model_Type <- ifelse(grepl("_LB$", p_values_and_impr$Model), "LB", "no_LB")

# View the updated dataframe
head(p_values_and_impr)
p_values_and_impr
```

@@

ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
```

@@

ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
```

@@

@@

@@

```
Total hospitalizations
